---
title: Salaries at Globe
output:
  github_document:
  html_preview: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  fig.width = 13,
  fig.height = 5
)
```

```{r}
library(ggplot2)
library(scales)
library(forcats)
library(tidytext)
library(dplyr)
library(xlsx)
```

### Read institute names and their abbreviations

```{r}
inst_df <- read.xlsx("salaries.xlsx", sheetName = "institutes", header = FALSE) %>% setNames("id")
parsed_ids <- strsplit(inst_df$id, " \\(|\\)")
inst_df <- inst_df %>%
  mutate(
    name = sapply(parsed_ids, `[[`, 1),
    id = sapply(parsed_ids, `[[`, 2)
  )

inst_df
```

### Read salary information

#### PhDs

```{r}
phd_df <- read.xlsx("salaries.xlsx", sheetName = "PhD") %>%
  .[-nrow(.), c("Institut", "Avg..Fuldtidsløn", "Median")] %>%
  setNames(c("inst_id", "mean", "median")) %>%
  mutate(role = "PhD")

phd_df
```

#### Postdocs

```{r}
postdoc_df <- read.xlsx("salaries.xlsx", sheetName = "postdoc") %>%
  .[-nrow(.), c("Institut", "Avg..Fuldtidsløn", "Median")] %>%
  setNames(c("inst_id", "mean", "median")) %>%
  mutate(role = "postdoc")

postdoc_df
```

#### Assistant professors

```{r}
adjunkt_df <- read.xlsx("salaries.xlsx", sheetName = "adjunkt") %>%
  .[-nrow(.), c("Institut", "Avg..Fuldtidsløn", "Median")] %>%
  setNames(c("inst_id", "mean", "median")) %>%
  mutate(role = "assistant professor")

adjunkt_df
```

#### Associate professors

```{r}
lektor_df <- read.xlsx("salaries.xlsx", sheetName = "lektor") %>%
  .[-nrow(.), c("Institut", "Avg..Fuldtidsløn", "Median")] %>%
  setNames(c("inst_id", "mean", "median")) %>%
  mutate(role = "associate professor")

lektor_df
```

#### Combine into a single tidy table

```{r}
salaries_df <- rbind(phd_df, postdoc_df, adjunkt_df, lektor_df) %>%
  mutate(
    median_year = median,
    median_month = median_year / 12,
    mean_year = mean,
    mean_month = mean_year / 12,
    is_globe = inst_id == "GLOBE",
    role = factor(role, levels = c("PhD", "postdoc",
                                   "assistant professor", "associate professor"))
  ) %>%
  as_tibble

salaries_df
```

#### Compute mean and median

```{r}
summary_df <- group_by(salaries_df, role) %>%
  summarise(
    mean_month = mean(mean_month),
    mean_year = mean(mean_year),
    median_month = median(median_month),
    median_year = median(median_year)
  )

summary_df
```

```{r}
plot_salaries <- function(salaries_df, summary_df, stat, time) {
  salaries_df_cols <- c("inst_id", "role", paste0(stat, "_", time), "is_globe")
  summary_df_cols <- c("role", paste0(stat, "_", time))

  sal_df <- salaries_df[salaries_df_cols]
  sum_df <- summary_df[summary_df_cols]
  
  salaries_df_cols2 <- c("inst_id", "role", "stat", "is_globe")
  summary_df_cols2 <- c("role", "stat")

  names(sal_df) <- salaries_df_cols2
  names(sum_df) <- summary_df_cols2

  sal_df %>%
  mutate(factor_id = reorder_within(inst_id, stat, within = role)) %>%
  ggplot(aes(factor_id, stat, color = is_globe)) +
    geom_point() +
    geom_hline(data = sum_df, aes(yintercept = stat, linetype = paste("SUND", !!stat), alpha = 0.8)) +
    scale_x_reordered() +
    scale_y_continuous(labels = label_comma(), breaks = breaks_pretty()) +
    scale_color_manual(values = c("black", "red")) +
    scale_linetype_manual(values = "dashed") +
    guides(color = "none", size = "none", alpha = "none", linetype = guide_legend("")) +
    labs(
      title = paste(tools::toTitleCase(stat), paste0(time, "ly"), "salaries across institutes at SUND"),
      x = "",
      y = "salary [dkk]"
    ) +
    theme_bw() +
    theme(text = element_text(size = 13),
          legend.position = "bottom",
          axis.text.x = element_text(hjust = 1, angle = 45)) +
    facet_wrap(~ role, scales = "free", nrow = 1)
}
```

#### Yearly statistics

```{r}
p_median_y <- plot_salaries(salaries_df, summary_df, "median", "year")
p_median_y

p_mean_y <- plot_salaries(salaries_df, summary_df, "mean", "year")
p_mean_y
```

#### Monthly statistics

These numbers are just the yearly numbers above divided by 12 (I personally find these
easier to understand than the yearly numbers, in the context of monthly rent and other
expenses, so I'm showing these as well).

```{r}
p_median_m <- plot_salaries(salaries_df, summary_df, "median", "month")
p_median_m

p_mean_m <- plot_salaries(salaries_df, summary_df, "mean", "month")
p_mean_m
```

#### Save all figures to a PDF

```{r}
pdf("salaries.pdf", width = 13, height = 5)
print(p_median_y)
print(p_mean_y)
print(p_median_m)
print(p_mean_m)
dev.off()
```

